%!TEX root = ../../report.tex
\newpage
\section{Hybrid approaches}
\label{sec:hybrid}
In sections \ref{sec:collaborative} and \ref{sec:content} we described collaborative and content-based recommender systems. These two have, of course, both advantages and disadvantages as also have described earlier. To overcome some of these disadvantages, one could combine the two approaches into a 'hybrid' recommender system.\newline

In the following a list of applied hybridized approaches will be presented, to outline some of the different methods for doing this.
According to \citet[p. 20-22]{TowardsTheNextGenerationOfRs} the methods can be classified as the following four: \textit{"(1) implementing collaborative and content-based methods separately and combining their predictions, (2) incorporating some content-based characteristics into a collaborative approach, (3) incorporating some collaborative characteristics into a content-based approach, and (4) constructing a general unifying model that incorporates both content-based and collaborative characteristics"}. It should be noted here that this list is not comprehensive and many other hybridized approaches exists. \newline
The first method is to implement the collaborative and content-based independently of each other and then obtain a recommendation for the same item from both, and then either combine the rating from each system into a single rating or select the 'best' rating relative to the recommended item based on a predetermined quality setting. \newline
The second method is to add some content-based characteristics to a collaborative recommender system. Here the system uses the profile learner to compare the similarity between users and thereby partly overcome some of the sparsity problems, because user-pairs most often does not have many ratings in common. Another feature of this approach is that users can be recommended an item based on, not only if it is highly rated by peer-users, but also if this item matches the users profile. \newline
The third method used some kind of dimensionality reduction technique on the profiles given from the 'profile learner' in the content-based recommender approach. In simple terms, this mean reducing the number of dimensions found in the vector space, in order to improve performance compared to the pure content-based approach. \newline
The fourth hybrid approach is the combination of the collaborative and content-based into one single recommendation model. The approach have been adopted by many researchers in recent years and because of this, many different methods of this hybridizing approach have been proposed \citep[p. 22]{TowardsTheNextGenerationOfRs}. These different methods somewhat lies outside of the scope of this report and will therefore not be further described. \newline

