%!TEX root = ../../report.tex
\section{Introduction}
In our everyday life we are constantly exposed to recommendations in some form or the other. Examples of this can be via televised commercials or signs in the supermarket.
It is possible for the television and supermarket advertisers to make more personal recommendations by the use of demographics such as statistics of which age group that watches tv during certain hours and who buys certain items in which geographical areas. However, from the group of people that are exposed to these recommendations, it is possible that only a fraction will actually find the commercial relevant to their inherent needs or interests. What if there existed a way to target each individual user with personal recommendations? While we might have to wait a little with that in regard to the television and supermarket \todo{(RFIDs are moving us closer to a reality in this area)slet?}, it does already exist on the Internet and other places and are known as: Recommender Systems.\newline
Where recommendations from the recommender systems described in this paper distinguish themselves from the above mentioned methods, is that the recommender systems are personalizing their recommendations for each of their individual users. This is done by aquiring information about each users preferences and interests, both given directly and indirectly by the user itself. Directly by the user typing in relevant information to the system and indirectly by observing the individual users behaviour via the trail of technological footprints that is left behind when the user interact with an individual website. These footprints includes items bought, items viewed, what persons with similar interests have bought, what they search for, etc. All this information is then used to compute, in various ways depending on which type of recommender system is used, the recommendations that can be presented to the user.

It is argued that we are moving away from an era of search based interaction and entering one of discovery\todo{argued hvor?}. The meaning  behind this idea is that we are moving away from knowing what we want and hence search for, and instead are shown new products or sites that we did not knew existed, or that we even needed. No matter what, recommender systems will continue to evolve and keep increasing their efficiency, possibly until they can tell more about you than you can yourself.


\todo{Reference this article: \url{http://money.cnn.com/magazines/fortune/fortune_archive/2006/11/27/8394347/index.htm}}

Quote:
"The effect of recommender systems will be one of the most important changes in the next decade," says University of Minnesota computer science professor John Riedl, who built one of the first recommendation engines in the mid-1990s. "The social web is going to be driven by these systems."

\todo{finish section}
\newpage

\subsection*{Research Question}
In the world today, all users experience recommendations on a daily basis, as suggested above. Recommender systems are embedded into services like Netflix, Amazon or LinkedIn and many other places. \newline
The purpose of this project is to get a deeper understanding of recommender systems and to possibly even suggest novel approaches in recommender systems design.\newline
The following questions will be discussed and possibility answers: How do Recommender Systems work? Do they work the same way or are different types of recommender systems in play? What role does a philisophical perspective play in the making of a recommender system? However, the main purpose of this report are to answer the following: \newline

\textbf{What defines a good recommender system and how can a recommender system for courses at ITU be made?}\newline

The answer to this question, will be found by following the structure presented below.

\subsection*{Overview}
This section are meant to provide the reader with a overview of the structure of the report, aswell as insight in the approach taken in the forecoming of the report.\newline
The report are overall split in four chapters, which also represents how the work process have been. The first chapter is about the different types of recommender system and how they work. This part describes two types of recommender systems and ways to make hybrid versions between the two.\newline
The second chapter represents the philosophical part of the project. This section describes a framework and certain values, that we have tried to used and embed in the recommender system for ITU. \newline
The third chapter reflects our conception of the recommender system for courses at ITU. This part of the report have been produced by using the obtained knowledge from chapter 1 and 2.
Chapter four contains a discussion and reflection of the proposed system and the project in general. This chapter concludes the report and project and does therefore provide the reader with some afterthoughts.\newline
\newpage